Optimised Cost Considering Huffman Code for Biological Data Compression

### 2015-01-28 Article/Article.tex
Abstract modified

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
\journal{Computers & Mathematics with Applications}
\bibliographystyle{elsarticle-num}
\begin{document}

@@

\title{Optimized and Cost Considering Huffman Code For Biological Data Transmission}

@@

\begin{abstract}
%% Text of abstract
In the recent year with the wide use of cloud and the distributed application in our life, the data transmission become more challenging over the years. In this paper we propose a new approach to improve the data transmission technique by improving the representation of Huffman code with the use of a genetic algorithm for optimal allocation of the codes to the symbols. The work in this research is consists of two steps. The First step of our algorithm is to generate a set of codes to the different symbols by cost considering of zero and ones. The aim of the codes generation method is to minimise the number of ones on the generated codes, which minimise the cost of the transmission. The second part of the approach is to find the optimal allocation of generated codes to different symbols, the optimal allocation is based on genetic algorithm to find the best allocation with a given penalty on cost, in order de sacrificing cost to minimise the number of bits. The case study of the approach is the biological data, which present a good application area to data compression with the huge growth of the biological data bases. The results are very promising and we have compared with the standard cost considering Huffman code and the original Huffman code.
\end{abstract}

@@

\FOR{each distinct symbol $i$} 
\STATE $ Enqueue\left(max\_Q , frequency~[~i~]\right)$
\ENDFOR
\STATE create a root node
\STATE $cost~[~root~]\leftarrow 0$
\STATE $Enqueue\left(min\_Q , cost~[~root~]\right)$
\STATE Define costs of the left and right child of the binary tree
\REPEAT 
\STATE $cost\_of\_parent\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE create $left$ and $right$ child for this node
\STATE $cost~[~left\_child~]\leftarrow cost\_of\_parent\_node+left\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~left\_child~]\right)$
\STATE $cost~[~right\_child~]\leftarrow cost\_of\_parent\_node+right\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~right\_child~]\right)$
\STATE Mark parent node as explored
\UNTIL{$2\left(n-1\right)$ nodes are created}
\WHILE{$min\_Q\neq\emptyset$} 
\STATE $leaf\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE $frequency[leaf\_node]\leftarrow Dequeue\left(max\_Q\right)$
\ENDWHILE
\FOR{each parent node $j$} 
\STATE $frequency~[~j~]\leftarrow frequency~[~left\_child~]+frequency~[~right\_child~]$
\ENDFOR
\REPEAT
\IF{conflict between nodes} 
\STATE resolve conflict by swapping conflicted nodes
\STATE calculate and reassign cost of all affected nodes
\STATE calculate and reassign frequency of all affected nodes 
\ENDIF
\UNTIL{all conflicts are resolved }
%\algstore{myalg}
\end{algorithmic}
\end{algorithm}
\subsection{Optimisation of the Codes}
\subsubsection{Problem formulation}
The problem of finding the best allocation of codes to each symbol can be   modelled as an Assignment Problems with Constraint, the problem is formulated as follows :\\\\
\textbf{Definition:} Given a set of codes $C=\left\{C_{1},C_{2}...C_{n}\right\}$, and a set of frequencies $C=\left\{Q_{1},Q_{2}...Q_{n}\right\}$. For each code we have the length of the code $|C_{i}|$ (number of bits) and the cost of the code $S_{C_{i}}$ (cost of ones and zeros), the objective is to assign to each frequency a code in order to minimize the total number of bits, while respecting the initial assignment total cost $S_{t}$ with a given penalty $\lambda$. This penalty coefficient represent the allowed Marge can be sacrificed for cost to optimize number of bits    
\\
The Objective Function is :

@@

Genetic Algorithm (GA) is a bio-inspired meta-heuristics algorithm developed by \cite{gen1}. GA is a stochastic optimization algorithm imitate the natural evolution process of genomes. GA started by generate a population of random feasible solutions. The optimization process of GA take this generate at each iteration a new population based on the previous one.  The process can be described as follow: Firstly the selection of two or more solution among the current population, by one of the well-known selection techniques \cite{bli95}. These selected solutions will be considered as a parents, the second operation of the genetic algorithm is the crossover, which take these parents as inputs to generate other new solutions considered as sons. The third operator of the genetic algorithm is the mutation which ensure a good diversification on the search process. The new solutions can be mutate according to a given mutation probability. The mutation change a position or more on the solution with new value different to the previous one. The quality of each solution is computed with the fitness function which control the evolution of the GA population by the deletion of the worst solution and insertion of the good solutions among  parents and sons. This processes is repeated until the stopped criteria is achieved which can be the number of generation or if the population is stabilized.

@@

The main objective of the GA optimisation algorithm for bits minimisation (GaBm) problem is to assign to each frequency a specific code. The encoding of an optimization problem solution into a chromosome is one of the most important issue to obtaining a good optimisation results. GaBM uses a two array fixed length to "64" which is the number of combination for all nucleotides, the first contain the frequencies of each codes and the second contain the cost of each codes. In this way, our genetic algorithm will work with the two array and uses the entry index on the allocation process.

@@

\begin{table}
\small
\label{datas}
\caption{influence of penalty on bit minimization}
\begin{center}
\begin{tabular}{c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Bits}$ &		\textbf{$\lambda (\%)$} \\\hline
Mycobacterium smegmatis &	 70760174 &37809636& 4\\\hline
Amycolatopsis benzoatilytica &	92010490& 49654132& 3\\\hline
Mycobacterium rhodesiae&	69421783& 37563017 &3\\\hline
Streptomyces bottropensis&	96294638& 50330594& 5\\\hline
Mycobacterium smegmatis. MC2&	74855668 &40380464& 5\\\hline
Mycobacterium smegmatis MKD8&	76738330& 40782100 &4\\\hline
Bradyrhizobium WSM471&	84667949 &45961149& 3\\\hline
Amycolatopsis thermoflava&	92416243 &48336727& 5\\\hline
Bacillus thuringiensis serovar thuringiensis&	66145783& 35330943& 4\\\hline
Bacillus thuringiensis Bt407&	68751359& 37306497 &3\\\hline
Pseudomonas aeruginosa 9BR&	72737300 &40264790& 2\\\hline
Bacillus thuringiensis serovar berliner&	67981988 &36770118 &3\\\hline
Bacillus thuringiensis serovar pakistani&	65896779& 35069999 &4\\\hline
Pseudomonas aeruginosa LES400&	71762305 &38210997 &4\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}
\begin{figure}[h]
\begin{center}
\includegraphics{Images/Drawing4.jpg}
\caption{Convergence of Optimized cost considering Huffman algorithm}
\end{center}
\label{Fig5}
\end{figure}
\section{Conclusion}
\label{sec5}
In this paper we have proposed a new approach for efficient data compression using Huffman code and optimised strategy. The new approach is divided into phases, firstly a cost considering Huffman algorithm are proposed which reduce the cost of the generated codes, these codes are secondary passed by the optimisation algorithm to reduce the global number of bits using a cost penalty.\\
The proposed approach is tested with biological genomic sequence and a performance comparison is made with the standard Huffman code and the cost considering without optimisation. Simulation results showed that the proposed approach is more robust and efficient compared to other competing algorithms because its penalty based optimisation strategy to search the best allocation of codes to different frequencies.

>>>>

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}
\journal{Computers \& Mathematics with Applications}
\bibliographystyle{elsarticle-num}
\begin{document}

@@

\title{Optimised Cost Considering Huffman Code for Biological Data Compression}

@@

\begin{abstract}
%% Text of abstract
Classical Huffman code has widely been used to compress biological datasets. Though a considerable reduction of size of data is obtained by classical Huffman code, yet, more efficient encoding is possible by treating binary bits differently considering requirement of transmission time, energy consumption, and likewise. A number of techniques already modified Huffman code algorithm to  obtain optimal prefix-codes for unequal letter costs in order to reduce overall transmission cost (time). 
%In the recent year with the wide use of cloud and the distributed application in our life, the data transmission become more challenging over the years. 
In this paper we propose a new approach to improve compression performance of one such extension, cost considering Huffman code,   %the data transmission technique by improving the representation of Huffman code with the use of a 
by applying genetic algorithm for optimal allocation of the codewords to the symbols. 
%The work in this research is consists of two steps. 
The approach start with generating a set of codewords for different input symbols based on their frequencies by considering cost (length) of 0 and 1 as $\alpha$ and $\beta$,~$\alpha < \beta$. This step ensures optimal cost encoding, however, compression performance is not higher than classical Huffman code. 
%The aim of the codes generation method is to minimise the number of ones on the generated codes, which minimise the cost of the transmission. 
To improve compression performance, the second part of the approach uses genetic algorithm to find an optimal allocation of generated codes to different symbols. The idea is to sacrifice some cost to minimise total number of bits, hence, the genetic algorithm works by giving penalty on cost. The performance of the approach is evaluated by applying it to compress some standard biological dataset and comparing the result with the results produced by classical Huffman code and standard cost considering Huffman code.
% The case study of the approach is the biological data, which present a good application area to data compression with the huge growth of the biological data bases. 
%The results are very promising and we have compared with the standard cost considering Huffman code and the original Huffman code.
\end{abstract}

@@

\FOR{each distinct symbol $i$} 
\STATE $ Enqueue\left(max\_Q , frequency~[~i~]\right)$
\ENDFOR
\STATE create a root node
\STATE $cost~[~root~]\leftarrow 0$
\STATE $Enqueue\left(min\_Q , cost~[~root~]\right)$
\STATE Define costs of the left and right child of the binary tree
\REPEAT 
\STATE $cost\_of\_parent\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE create $left$ and $right$ child for this node
\STATE $cost~[~left\_child~]\leftarrow cost\_of\_parent\_node+left\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~left\_child~]\right)$
\STATE $cost~[~right\_child~]\leftarrow cost\_of\_parent\_node+right\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~right\_child~]\right)$
\STATE Mark parent node as explored
\UNTIL{$2\left(n-1\right)$ nodes are created}
\WHILE{$min\_Q\neq\emptyset$} 
\STATE $leaf\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE $frequency[leaf\_node]\leftarrow Dequeue\left(max\_Q\right)$
\ENDWHILE
\FOR{each parent node $j$} 
\STATE $frequency~[~j~]\leftarrow frequency~[~left\_child~]+frequency~[~right\_child~]$
\ENDFOR
\REPEAT
\IF{conflict between nodes} 
\STATE resolve conflict by swapping conflicted nodes
\STATE calculate and reassign cost of all affected nodes
\STATE calculate and reassign frequency of all affected nodes 
\ENDIF
\UNTIL{all conflicts are resolved }
%\algstore{myalg}
\end{algorithmic}
\end{algorithm}
\subsection{Optimisation of the Codes}
\subsubsection{Problem formulation}
The problem of finding the best allocation of codes to each symbol can be   modelled as an Assignment Problems with Constraint, the problem is formulated as follows :\\\\
\textbf{Definition:} Given a set of codes $C=\left\{C_{1},C_{2}...C_{n}\right\}$, and a set of frequencies $C=\left\{Q_{1},Q_{2}...Q_{n}\right\}$. For each code we have the length of the code $|C_{i}|$ (number of bits) and the cost of the code $S_{C_{i}}$ (cost of ones and zeros), the objective is to assign to each frequency a code in order to minimize the total number of bits, while respecting the initial assignment total cost $S_{t}$ with a given penalty $\lambda$. This penalty coefficient represent the allowed Marge can be sacrificed for cost to optimise number of bits    
\\
The Objective Function is :

@@

Genetic Algorithm (GA) is a bio-inspired meta-heuristics algorithm developed by \cite{gen1}. GA is a stochastic optimisation algorithm imitate the natural evolution process of genomes. GA started by generate a population of random feasible solutions. The optimisation process of GA take this generate at each iteration a new population based on the previous one.  The process can be described as follow: Firstly the selection of two or more solution among the current population, by one of the well-known selection techniques \cite{bli95}. These selected solutions will be considered as a parents, the second operation of the genetic algorithm is the crossover, which take these parents as inputs to generate other new solutions considered as sons. The third operator of the genetic algorithm is the mutation which ensure a good diversification on the search process. The new solutions can be mutate according to a given mutation probability. The mutation change a position or more on the solution with new value different to the previous one. The quality of each solution is computed with the fitness function which control the evolution of the GA population by the deletion of the worst solution and insertion of the good solutions among  parents and sons. This processes is repeated until the stopped criteria is achieved which can be the number of generation or if the population is stabilized.

@@

The main objective of the GA optimisation algorithm for bits minimisation (GaBm) problem is to assign to each frequency a specific code. The encoding of an optimisation problem solution into a chromosome is one of the most important issue to obtaining a good optimisation results. GaBM uses a two array fixed length to "64" which is the number of combination for all nucleotides, the first contain the frequencies of each codes and the second contain the cost of each codes. In this way, our genetic algorithm will work with the two array and uses the entry index on the allocation process.

@@

\begin{table}
\small
\label{datas}
\caption{influence of penalty on bit minimization}
\begin{center}
\begin{tabular}{c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Bits}$ &		\textbf{$\lambda (\%)$} \\\hline
Mycobacterium smegmatis &	 70760174 &37809636& 4\\\hline
Amycolatopsis benzoatilytica &	92010490& 49654132& 3\\\hline
Mycobacterium rhodesiae&	69421783& 37563017 &3\\\hline
Streptomyces bottropensis&	96294638& 50330594& 5\\\hline
Mycobacterium smegmatis. MC2&	74855668 &40380464& 5\\\hline
Mycobacterium smegmatis MKD8&	76738330& 40782100 &4\\\hline
Bradyrhizobium WSM471&	84667949 &45961149& 3\\\hline
Amycolatopsis thermoflava&	92416243 &48336727& 5\\\hline
Bacillus thuringiensis serovar thuringiensis&	66145783& 35330943& 4\\\hline
Bacillus thuringiensis Bt407&	68751359& 37306497 &3\\\hline
Pseudomonas aeruginosa 9BR&	72737300 &40264790& 2\\\hline
Bacillus thuringiensis serovar berliner&	67981988 &36770118 &3\\\hline
Bacillus thuringiensis serovar pakistani&	65896779& 35069999 &4\\\hline
Pseudomonas aeruginosa LES400&	71762305 &38210997 &4\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}
\begin{figure}[h]
\begin{center}
\includegraphics{Images/Drawing4.jpg}
\caption{Convergence of Optimised cost considering Huffman algorithm}
\end{center}
\label{Fig5}
\end{figure}
\section{Conclusion}
\label{sec5}
In this paper we have proposed a new approach for efficient data compression using Huffman code and optimised strategy. The new approach is divided into phases, firstly a cost considering Huffman algorithm are proposed which reduce the cost of the generated codes, these codes are secondary passed by the optimisation algorithm to reduce the global number of bits using a cost penalty.\\
The proposed approach is tested with biological genomic sequence and a performance comparison is made with the standard Huffman code and the cost considering without optimisation. Simulation results showed that the proposed approach is more robust and efficient compared to other competing algorithms because its penalty based optimisation strategy to search the best allocation of codes to different frequencies.
